DeepPicar: A Low-cost Deep Neural Network-based Autonomous Car

### 2018-02-01 paper/intro.tex
intro revision

% Findings:
%
% - we find DNN processing is an ideal real-time workload from the
%   perspective of timing predictability as the amount of computation
%   needed is fixed at design time and doesn't change over input.
%
% - real-time processing of non-trivial DNN inferencing is feasible
%   on today's embedded multicore platform, even as inexpensive as
%   raspberry pi3.
%
% - all tested platforms have enough computing capacity to consolidate 
%   multiple DeepPicar's CNN workloads or other compute intensive tasks
%   simultaneously.
%
% - however, when consolidating multiple workloads shared resource
%   contention is an important issue. we carefully analyzed
%   sensitivity of DeepPicar's workload with respect to shared
%   hardware resources.
%
% - in particular, we find DNN processing is highly sensitive to
%   memory performance. performance severely affected when memory
%   bandwidth is contented.
%
% - however, we find that dnn processing is not sensitive to cache
%   space. we find applying cache partitioning does not help improve
%   isolation of DNN workload when memory bandwidth is contended.
From the systematic study, we made a number of interesting
observations from the perspective of real-time systems.
% predictability
First, we find that DNN inferencing is, in a sense, an ideal real-time
workload as the amount of computation needed to compete a single
inference is fixed at the DNN architecture design time and does not
change at runtime over different inputs (e.g., different image frame).
This means that it is feasible to achieve highly predictable and
tightly bounded worst-case execution time estimation.

% performance on low-cost embedded multicore
Second, we find that real-time processing of DeepPicar's DNN-based
control loop is feasible on today's embedded computing platforms, even 
as inexpensive as the Raspberry Pi 3. We find that the control loop 
completes in under 33.$\overline{\mbox{33}}$ ms, or 30 Hz, using just two 
cores of the Raspberry Pi 3's quad-core CPU (w/o using its GPU), and can
do so 100\% of the time; we can achieve 20 Hz control frequency using
just one core of the Raspberry Pi 3's quad-core CPU.
Other tested embedded platforms, Intel UP and NVIDIA TX2, offer even
better performance, and are capable of supporting deep-learning based 
real-time control over 100 Hz control frequency on the TX2 when its
embedded GPU is used.
We find that all the tested embedded computing platforms have 
enough computing capacity to consolidate multiple instances of
DeepPicar's DNN workload.

% all platforms have enough capacity to consolidate
Third, contention in shared hardware resources remains an important
issue when consolidating multiple workloads. 
In particular, we find that DNN inferencing is highly sensitive to
memory performance inference as we observe up to an 11.6X control loop
execution time increase, when memory performance intensive
applications are co-scheduled on idle CPU cores.
We find, however, that DNN inferencing is not sensitive to cache
space; we show that applying cache partitioning has negligible impact
in providing timing isolation on consolidated scenarios. This finding
suggests that in order to guarantee real-time performance of DNN-based
control applications, solutions to guarantee its memory performance
are much more important.

% contributions
The {\bf contributions} of this paper are as follows:
\begin{itemize}
  \item We provide the design and implementation of a low-cost
    autonomous car testbed, DeepPicar, which is based on a
    state-of-the-art end-to-end deep learning, as
    open-source~\footnote{The source code, build 
instruction the DeepPicar
can be found at: \url{https://github.com/heechul/picar}}, with the
    goal of lowering the economic and safety-related barriers to study
    autonomous cars.

  \item We systematically analyze the characteristics of DeepPicar's
    DNN inferencing workload on a number of representative
    contemporary embedded computing platforms from the perspective of
    real-time systems. In particular, to the best of our knowledge,
    our work is first to investigate the impacts of shared resource
    contention on DNN inferencing workloads in the context of
    real-time systems.

@@

The remaining sections of the paper are as follows: Section II 
provides a background of applications in autonomous driving and related works.
 Section III gives an overview of the DeepPicar platform, including the 
high-level system and the methods used for training and inference. 
Section IV presents our evaluation of the platform and how different 
factors can affect performance. Section V gives a comparison between 
the Raspberry Pi 3 and other embedded computing platforms to 
determine their suitably for autonomous driving research. The paper finishes with 
conclusions in Section VI.

>>>>

% Findings:
%
% - we find DNN processing is an ideal real-time workload from the
%   perspective of timing predictability as the amount of computation
%   needed is fixed at design time and doesn't change over input.
%
% - real-time processing of non-trivial DNN inferencing is feasible
%   on today's embedded multicore platform, even as inexpensive as
%   raspberry pi3.
%
% - all tested platforms have enough computing capacity to consolidate 
%   multiple DeepPicar's CNN workloads or other compute intensive tasks
%   simultaneously.
%
% - however, when consolidating multiple workloads shared resource
%   contention is an important issue. we carefully analyzed
%   sensitivity of DeepPicar's workload with respect to shared
%   hardware resources.
%
% - in particular, we find DNN processing is highly sensitive to
%   memory performance. performance severely affected when memory
%   bandwidth is contented.
%
% - however, we find that dnn processing is not sensitive to cache
%   space. we find applying cache partitioning does not help improve
%   isolation of DNN workload when memory bandwidth is contended.
From the systematic study, we made a number of interesting
observations from the perspective of real-time systems.
% predictability
First, we find that DNN inferencing is highly predictable---from the
timing perspective---as the amount of computation needed to complete a
single inference is fixed at the DNN architecture design time and does
not change at runtime over different inputs (e.g., different image
frames). This predictable timing behavior is obviously a desirable
property for real-time systems.
%% , which gives another reason for the
%% real-time community to be interested in the use of DNNs.

% performance on low-cost embedded multicore
Second, we find that real-time processing of DeepPicar's CNN-based
control loop is feasible on today's embedded computing platforms, even 
as inexpensive as the Raspberry Pi 3. We find that the control loop 
completes in under 33.$\overline{\mbox{33}}$ ms, or 30 Hz, using just two 
cores of the Raspberry Pi 3's quad-core CPU (w/o using its GPU), and can
do so 100\% of the time; or we can achieve 20 Hz control frequency using
just one core of the Raspberry Pi 3's quad-core CPU.
Other tested embedded platforms, Intel UP and NVIDIA TX2, offer even
better performance, and are capable of supporting deep-learning based 
real-time control over 100 Hz control frequency on the TX2 when its
embedded GPU is used.
We find that all the tested embedded computing platforms have 
enough computing capacity to consolidate multiple instances of
DeepPicar's DNN workload.

% all platforms have enough capacity to consolidate
Third, contention in shared hardware resources remains an important
issue when consolidating multiple workloads. 
In particular, we find that DNN inferencing is highly sensitive to
memory performance interference as we observe up to an 11.6X control
loop execution time increase, when memory performance intensive
applications are co-scheduled on idle CPU cores.
On the other hand, we find that DNN inferencing is not sensitive to
cache space.
%% ; we show that applying cache partitioning has negligible impact
%% in providing timing isolation on consolidated scenarios.
This finding
suggests that in order to guarantee real-time performance of DNN-based
control applications, solutions to guarantee their memory performance
are more important.

% contributions
The {\bf contributions} of this paper are as follows:
\begin{itemize}
  \item We present DeepPicar, a low-cost autonomous car testbed, which
    is based on a state-of-the-art end-to-end deep learning, as 
    open-source~\footnote{The source code, build 
instruction the DeepPicar
can be found at: \url{https://github.com/heechul/picar}}, with the
    goal of lowering the economic and safety-related barriers to study
    autonomous cars.

  \item We systematically analyze real-time characteristics of DeepPicar's
    DNN inferencing workload on a number of representative
    contemporary embedded computing platforms. To the
    best of our knowledge, no prior work investigated the
    impacts of shared resource contention on DNN inferencing workloads
    in the context of real-time systems.

@@

The remaining sections of the paper are as follows.
Section~\ref{sec:background}
provides a background of applications in autonomous driving and related works.
Section~\ref{sec:overview} gives an overview of the DeepPicar
platform, including the high-level system and the methods used for
training and inference. Section~\ref{sec:evaluation} presents our
evaluation of the platform and how different factors can affect
performance. Section~\ref{sec:comparison} gives a comparison between  
the Raspberry Pi 3 and other embedded computing platforms to 
determine their suitably for autonomous driving research.
We discuss more detailed training methods and system-level issues
in~\ref{sec:discussion}. We review related work in
Section~\ref{sec:related} and conclude in
Section~\ref{sec:conclusion}.
